\section{Consensus Algorithms}
In this world where the network can not be trusted, time lies to us and servers will randomly crash and burn how can we get anything done at all? Lets discuss how we can build a system we can trust, a system that behaves \textit{consistantly}. To build such a system we need the parts that make up the system to agree with eachother, the must have \emph{Consensus}. Here I discuss three well known solutions. Before we get to that lets look at the principle that underlies them all: \emph{The truth is defined by the majority}.

\subsubsection*{Quorums}
Imagine a node hard at work processing requests from its siblings, suddently it gets pre-empted. The other nodes notice it is no longer responding and declear it dead, they dont know its threads got paused. A few seconds later the node responds again as if nothing had happend, and in truth, unless it checks the system clock, from its perspective no time has paused. Alternatively a network error might partition the system, each group of servers can reach eachoter but not the others. The nodes in the group will declear those in the other group dead and continue their work. Usually this results in data loss if the work progresses at all.

We can solve this by voting over each descision. It will be a strange vote, no node cares about the descision itself. In most implementations the nodes only checks if it regards the sender as trustwothy or alive and then vote yes. To prove liveliness the vote proposal could include a number. Voters only vote yes if the number is correct. For example if the number is the highest they have seen. If a majority votes yes the node that requested the vote can be sure its not dead or disconnected. This is the idea behind "Quorums," majorities of nodes that vote.

\subsubsection*{paxos}
The \textsc{Paxos} algorithm\cite{paxos} uses a quorum to provide concensus. It does this by chosing a value among proposals such that only that value can be read as the accepted value. Usually it is used to build a fault tolerant distributed state machine. 

In \textsc{Paxos} there are three roles: proposer, acceptor and learner. It is possible for nodes to fullfil only one or two of these roles. For the rest of this explanation assume each node fullfils all three. To reach consensus on a new value we go through two phases: prepare and accept. Once the majority of the nodes has accepted a proposal the vale of that proposal been chosen. In \textsc{Paxos} nodes keep track of the highest proposal number $n$ they have seen. Lets go through a \textsc{Paxos} iteration from the perspective of a node trying to share something, \textit{a value}.

In the first phase a new \textit{value} is proposed by our node. It sends a \textit{perpare} request to a majority of acceptors. The request contains a proposal number $n$ higher then the highest number our node has seen up till now. The number is unique to our node. Each acceptor only responds if our number $n$ is the highest it has seen. If an acceptor had already accepted one or more requests it includes the accepted proposal with the highest $n$ in its respons.

In phase two our node checks if it got a response from the majority. Our node is going to send an accept request back to those nodes. The content of the accept request depends on what our node recieved in response to its prepare request:
%
\begin{enumerate}
	\item Our node recieved a response with number $n_p$. This means an acceptor has already accepted a value. If we continued with our own value the system would have two different accepted values. Therefor the content of our accept request will be the value from proposal $n_p$.
	\item It recieved only acknowleding replies and none contained a previously accepted value. The system has not yet decided on a value. The content of our accept request will be the value or node wants to propose but with our number $n$.
\end{enumerate}
%
The acceptors accept the request if they did not yet recieve a prepare request numberd greater then $n$. On accepting a request an acceptor sends a message to all learners\footnote{remember usually every node is a learner}. This way the learners learn a new value as soon as its ready.

% Lets get a feeling why this works by looking at what happens during network failure. Imagine the case where an accept with value $v_a$ is send to the majority $m$. The network fails and only $m-1$ nodes accept. Now 50\% of the nodes will reply $v_a$ as chosen value to learners. Any majority response to a propose request will include a node from $m-1$, thus with the accepted value. No accept request with another value then $v_a$ can therefore be send. Another value $v_b$ will thus never be accepted.

Lets get a feeling why this works by looking at what happens during node failure. Imagine a case where a minimal majority $m$ accept value $v_a$. A single node in $m$ freezes after the first learners learned of the now chosen value $v_a$. After freezing $m-1$ of the nodes will reply $v_a$ as value to learners. The learners will conclude no value has been chosen given $m-1$ is not a majority\footnote{this is not yet inconsistent, \textit{Paxos} does not guarentee consistency over wether a value has been chosen}. As seen above acceptors change their value if they recieve a higher numberd accept request. If a single node changes its value to $v_b$ consensus will break since $v_a$ has already been seen as the chosen value by a learner. A new proposal that can result into higher numberd accept requests needs a majority response. A majority response will include a node from $m-1$. That node will incude $v_a$ as the accepted value. The value for the accept request changes to $v_a$. No accept request with another value then $v_a$ can thus be issued. Another value $v_b$ will therefore never be accepted. The new accept by issued to a majority changing at least one node to have accepted $v_a$. Now at least $m+1$ nodes have $v_a$ as accepted value.



Three roles proposer, acceptor and learner
Two phases, prepare and accept
 	key insight: an accepted prepare with a higher number will spread by being the awnser to new prepares
Usually build upon with an distributed state machine 
	a leader is elected and will handle all proposals
	leader recieves client requests and turns them into state machine commands
	these are proposed, when accepted they are executed (applied) to the current state
	


\subsubsection*{raft}
\subsubsection*{consensus as a service}
% zookeeper
